---
title: "CUBE - UOttawa"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_float: true
    theme: spacelab
    highlight: kate
---

Manuscript draft (google doc) available here:   https://docs.google.com/document/d/15joiEhDGsYJTgS0y5KkjgVT_Usv1emeCA80fxc0Odkk/edit?usp=sharing

----

## TODO

- <input type="checkbox" checked> Site summary data</input>
- <input type="checkbox" checked> Biweekly summary of variables</input>
  - <input type="checkbox" checked> Figures plus events timeline</input>
  - <input type="checkbox" checked> Correlation & Heatmap</input>
  - <input type="checkbox" checked> Add case counts / fix figure</input>
  
- <input type="checkbox" checked> Case Date Imputation. about that..</input>

- <input type="checkbox" checked>  Model for cases ~ .</input>
- <input type="checkbox" checked>  Test high vs low traffic as FE in RE model</input>
- <input type="checkbox" checked>  Finalize fig. 2</input>
- <input type="checkbox" checked>  Finalize fig. 3</input>
- <input type="checkbox" unchecked>  Fig. 4: refresh data </input>
- <input type="checkbox" unchecked>  Finish written methods and results</input>

----

<details>
<summary>Analysis code</summary>

```{r setup, echo=F}
knitr::opts_chunk$set(echo = T, message = F, warning = F, fig.align = 'center')
```

```{r data analysis, echo=TRUE}
# setup ---------------------------------------------
library(tidyverse)
library(here)
library(ggiraph)
library(patchwork)
library(ggtext)
library(janitor)

ggplot2::theme_set(theme_bw() + theme(
  legend.text = element_markdown(family = 'IBM Plex Sans', colour = 'gray30'),
  axis.text = element_markdown(family = 'IBM Plex Sans', colour = 'gray30'),
  axis.title = element_markdown(family = 'IBM Plex Sans',colour = 'gray30'),
))
hrbrthemes::import_plex_sans()

source(here('R/report_plots.R'))

# Functions -----------------------------------------

convert_cq_to_copies <- function(cq) 10 ** ((40.1 - cq) / 3.23)

# creates yyyy-ww label for grouping data
get_date_week <- function(x){
  y <- lubridate::year(x)
  w <- lubridate::week(x) |> str_pad(2, 'left', 0)
  str_glue("{y}-{w}")
}

# reverses get_date_week to Weds date during week
week_to_date <- function(year_week){
  year <- str_extract(year_week, '^....')
  d <- str_glue('{year}-01-01') |> as_date()
  wk <- str_extract(year_week, '..$') |> as.integer()
  ydays <- if_else(
    year == '2021',
    days(7*wk - 2),
    days(7*wk - 3),
  )
  return(d + ydays)
}


# get biweekly bin date for vector of dates
get_date_biweekly <- function(x){
  day_of_week <-  lubridate::wday(x)
  case_when(
    day_of_week %in% 1:3 ~ x + 3 - day_of_week,
    TRUE ~ x + 5 - day_of_week,
  )
}

# dates 2021-2022 with biweekly groups
biweekly_date_sequence <- function(){
  tibble(
    date = seq(
      as_date('2021-01-01'), 
      as_date('2022-12-31'), 
      by = 1)
  ) |> 
    mutate(biweek = get_date_biweekly(date))
}

# lookup table for date
week_midpoint_date_lookup <- function(start = '2021-01-01', end = '2023-01-01'){
  ends = lubridate::as_date(c(start, end))
  tibble(
    date = seq(ends[1], ends[2], by = 1),
    date_week = lubridate::week(date) |> str_pad(2, 'left', '0'),
    date_year = lubridate::year(date),
    week = str_glue("{date_year}-{date_week}"),
  ) |> 
    group_by(week) |> 
    summarise(date = mean(date))
}

# convert site1/site2 to hi/low traffic
get_traffic_level <- function(location){
  if_else(
    str_detect(location, 'Site 1'), 
    'high traffic', 
    'low traffic'
  )
} 

binom_ci <- function(x, n){
  Hmisc::binconf(x, n, method = 'wilson', alpha = 0.05) |> 
    as_tibble() |> 
    janitor::clean_names()
}

get_binom_ci <- function(data){
  data |> 
    summarise(
      x = sum(pcr_positive == 'Positive'),
      n = n(),
      binconf = map2(x, n, ~binom_ci(.x, .y))
    ) |> 
    unnest(binconf) |> 
    mutate(across(c(point_est, lower, upper), ~(.x*100) |> round(1)))
}

# Plot Elements ---------------------------------------

theme_color <- 'cornflowerblue'

# layout x-axis
scale_x_study_dates <- function(){
    scale_x_date(
      date_breaks = 'month',
      date_labels = month.name[c(9:12, 1:5)] |> strtrim(3) |> 
        paste0(' ', c(rep(2021, 4), rep(2022, 5))),
      limits = c(
        min(swabs_tidy$date_swab), 
        max(swabs_tidy$date_swab)
      ))
}
# get rid of x-axis for vertically stacked plots
theme_no_x_labels <- function(){
  theme(
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.text.x = element_blank()
  )
}
# get rid of y-axis for horizontally stacked plots
theme_no_y_labels <- function(){
  theme(
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank()
  )
}
# no minor grid and adjust spacing for multipanel
theme_remove_grid <- function(){
  theme(
    axis.title.y = element_blank(),
    panel.grid.minor = element_blank(),
    plot.title.position = 'panel', 
    plot.title = element_markdown(
      vjust = -1, 
      size = 8, 
      face = 'bold', 
      family = 'IBM Plex Sans',
      colour = 'gray50',
      margin = margin(0, 0, 0, 0)),
    plot.margin = unit(c(0, 0, 0, 0), 'mm'))
}

# Data ------------------------------------------------

# swab data
swabs <- 
  read_rds(here('data/cube.rds')) |> 
  filter(str_detect(site, '^UO_')) 

# filtered swabs without controls or sponge samples
swabs_tidy <- swabs |> 
  filter(!negative_control, 
         swab_type != 'sponge',
         date_swab < '2022-04-27') |> 
  select(date_swab, site, floor, location, pcr_positive, pcr_ct, co2) |> 
  mutate(biweek = get_date_biweekly(date_swab))

# lookup table for waste water site names
lookup_ww <- tribble(
  ~site_abbr, ~site,
  'UO_AA',	'Annex Residence',
  'UO_FA',	'Faculty of Social Sciences',
  'UO_FT',	'Friel Residence',
  'UO_NA',	'Northern sampling site',
  'UO_RGN',	'Roger Guindon Hall',
  'UO_ST',	'Southern sampling site',
  'UO_TBT',	'Tabaret Hall'
)

# UOttawa waste water data from R. Delatolla
uottawa_ww <-
  readxl::read_excel(here::here('data/ww_university.xlsx')) |>
  janitor::clean_names() |> 
  mutate(sample_date = as_date(sample_date)) |> 
  filter(
    sample_date > '2021-09-15', 
    sample_date < '2022-05-05',
    !is.na(viral_copies_per_copies_pep_avg),
    site != 'UO_RGN'
    ) |> 
  left_join(lookup_ww, by = c('site' = 'site_abbr')) |> 
  mutate(
    source =
      source_name |>
      str_remove('^Data - uOttawa - ') |>
      str_remove('.xlsx$')
  ) |> 
  transmute(site,
            sample_date, 
            signal = viral_copies_per_copies_pep_avg)

# ottawa wastewater data: daily means
regional_ww <- 
  read_rds(here('data/ww_ottawa.rds')) |> 
  select(sample_date, starts_with('cov_')) |> 
  pivot_longer(contains('cov_')) |> 
  mutate(target = str_extract(name, 'cov_n.'),
         stat = str_extract(name, 'mean|sd'),
         ) |> 
  select(-name) |> 
  pivot_wider(names_from = stat, values_from = value) |> 
  mutate(.lower = mean - sd, .upper = mean + sd)   

# data from university of ottawa
wifi <- 
  read_rds(here('data/uo_wifi.rds')) |> 
  filter(date >= min(swabs$date_swab), 
         date <= max(swabs$date_swab))

# important dates for context
event_dates <- 
  tribble(
    ~event, ~start, ~end,
    'Reading Week',  '2021-10-24', '2021-10-30',
    'Holiday Break', '2021-12-22', '2022-01-04',
    'Closure',       '2022-01-04', '2022-01-31',
    'Closure',       '2022-02-16', '2022-02-21',
    'Reading Week',  '2022-02-20', '2022-02-26',
  ) |> 
  mutate(across(start:end, as_date))

# UOttawa cases data
# clean data from: source(here('R/01_impute_missing_case_dates.R'))
cases <-
  read_rds(here('data/cases_rule_imputed.rds')) |> 
  select(case, role, case_date, UC:TBT) |>
  mutate(biweek = get_date_biweekly(case_date)) |> 
  nest(associated_sites = UC:TBT)


total_n_cases_study_period <- 
  cases |> 
  filter(case_date > '2021-09-20') |> 
  nrow()


# Summaries -----------------------------------------
# 
positivity_summary <- function(swabs){
  swabs |> 
    summarise(
      n = n(),
      x = sum(pcr_positive == 'Positive', na.rm = F),
      positivity = round(100 * x / n, digits = 1),
      .groups = 'drop')
}

co2_summary <- function(swabs){
  swabs |> 
    summarise(
      n = n(),
      co2_vals = list(co2),
      co2_mean = mean(co2, na.rm = T),
      .groups = 'drop')
}
  
# overall summary
swab_summary <- swabs_tidy |> positivity_summary()

# site summary
swab_summary_sites <- swabs_tidy |> 
  group_by(site) |> 
  positivity_summary()
  
# high-low traffic summary
swab_summary_location_traffic <- 
  swabs_tidy |> 
  mutate(traffic = get_traffic_level(location)) |> 
  group_by(traffic) |> 
  positivity_summary()

## Figure 1: data aggregation --------------------------

# uottawa cases - biweekly counts 
cases_biweekly <- 
  cases |> 
  select(case, biweek, role) |> 
  group_by(biweek) |> 
  summarise(cases = n(),
            cases_student = sum(role == 'Student'),
            cases_staff = sum(role == 'Employee'),
            ) |> 
  right_join(
    biweekly_date_sequence() |> 
      filter(biweek < '2022-02-03') |> 
      distinct(biweek),
    by = 'biweek'
  ) |> 
  mutate(across(where(is.integer), ~if_else(is.na(.), 0L, .))) 

# swabs biweekly summary
swabs_biweekly <- 
  swabs_tidy |>
  group_by(biweek) |> 
  positivity_summary() |> 
  left_join(swabs_tidy |> group_by(biweek) |> co2_summary())

# swabs biweekly summary by site  
swabs_biweekly_by_site <- 
  swabs_tidy |>
  group_by(site, biweek) |> 
  positivity_summary() |> 
  left_join(swabs_tidy |> group_by(site, biweek) |> co2_summary())

# UOttawa ww biweekly means
uottawa_ww_biweekly <- 
  uottawa_ww |> 
  mutate(biweek = get_date_biweekly(sample_date)) |> 
  group_by(biweek) |> 
  summarise(signal = mean(signal),
            n = n())

# daily ww data for study period
regional_ww_daily <- 
  regional_ww |> 
  filter(
    sample_date >= min(swabs_tidy$date_swab) - 4,
    sample_date <= max(swabs_tidy$date_swab) + 4,
  ) |> 
  group_by(sample_date) |> 
  summarise(mean = mean(mean))

# regional ww biweekly means
regional_ww_biweekly <- 
  regional_ww_daily |> 
  mutate(biweek = get_date_biweekly(sample_date)) |> 
  group_by(biweek) |> 
  summarise(mean = mean(mean))

# wifi site agg
wifi_extended <- read_rds(here('data/uo_wifi.rds')) |> 
  filter(date >= min(swabs$date_swab) - 2, 
         date <= max(swabs$date_swab) + 2) |> 
  mutate(biweek = get_date_biweekly(date))

# overall wifi biweekly timeseries
wifi_biweekly <- 
  wifi_extended |> 
  group_by(biweek) |> 
  summarise(
    n = n(),
    min = min(clients, na.rm = T),
    mean = mean(clients, na.rm = T),
    max = max(clients, na.rm = T)
  )
# per building wifi biweekly timeseries
wifi_biweek_sites <- 
  wifi_extended |> 
  group_by(biweek, site) |> 
  summarise(
    n = n(),
    min = min(clients),
    mean = mean(clients),
    max = max(clients)
  )

# Map --------------------------------------------------

swab_coords <- tribble(
  ~site, ~name, ~lat, ~lng,
  'MRT', 'Morrisette Hall (MRT)', 45.4232391, -75.684289,
  'MNT', 'Montpetit', 45.4225417102, -75.6826587146,
  '90U', '90 University', 45.422425557, -75.68501449,
  'DMS', 'Desmarais Bldg.', 45.4238767934, -75.687270754,
  'TBT', 'Tabaret Hall', 45.4245094016, -75.6863190018,
  'LPR',  'Louis Pasteur Bldg.', 45.42137566861, -75.6802638999,
)

ww_coords <- tribble(
  ~site, ~name, ~lat, ~lng,
  'aa', 'Annex Residence', 45.4267812902, -75.6803440596,
  'fa', 'Faculty of Soc. Sci.', 45.4216247, -75.6839038601,
  'ft', 'Friel Residence', 45.4304344008, -75.6829076653,
  'tbt', 'Tabaret Hall', 45.4245094016, -75.6863190018
  # 'na', 'North sampling Site', NA, NA,
  # 'st', 'South Sampling Site', NA, NA, # Nobody knows where this is
)

figure_sites_map <- 
  leaflet::leaflet(swab_coords, width = 600, height = 330) |> 
  leaflet::addProviderTiles(leaflet::providers$CartoDB.Positron) |> 
  leaflet::setView(lng = -75.6843, lat = 45.4235, zoom = 14) |> 
  leaflet::addCircleMarkers(~lng, ~lat, label =  ~name, popup = ~name, 
                            radius = 2, color = '#440099') |> 
  leaflet::addLabelOnlyMarkers(
    ~lng, ~lat, label =  ~site, 
    labelOptions = leaflet::labelOptions(
      noHide = T, direction = 'top', textOnly = T, 
    )) |> 
  leaflet::addCircleMarkers(~lng, ~lat, label =  ~site, 
                            radius = 2, color = '#440099') |> 
  leaflet::addCircleMarkers(
    data = ww_coords, ~lng, ~lat, label =  ~name, popup =  ~name, 
    radius = 4, color = '#f96999')


# Results ----------------------------------------------------

rs_data <- lst(
  swabs_collected = nrow(swabs_tidy),
  positivity_ovr = swab_summary$positivity,
  positivity_site_min = min(swab_summary_sites$positivity),
  positivity_site_max = max(swab_summary_sites$positivity),
)

rs_abstract <- rs_data |> 
  glue::glue_data(
    "Over the 32-week study period, we collected {swabs_collected} swabs at six university buildings, including two buildings where wastewater surveillance was currently ongoing. 
Overall, {positivity_ovr}% of swabs were PCR-positive for SARS-CoV-2, with individual buildings ranging between {positivity_site_min}% and {positivity_site_max}%.
*Comment on when spikes in cases, swabs, and waste-water occurred….* *Comment on prediction of cases using swabs, ww, or combined approach…*
"
  )

# rs_results <- rs_data |> glue::glue_data()


## corr -----

# joined data for pairwise correlation 
biweekly_data <- swabs_biweekly |> 
  transmute(biweek, `Swab PCR` = positivity, `CO2` = co2_mean) |> 
  left_join(
    wifi_biweekly |> transmute(biweek, `Wi-Fi` = mean),
    by = 'biweek'
    ) |> 
  left_join(
    uottawa_ww_biweekly |> transmute(biweek, `University WW` = signal),
    by = 'biweek'
  ) |> 
  left_join(
    regional_ww_biweekly |> transmute(biweek, `Ottawa WW` = mean),
    by = 'biweek'
  ) |> 
  left_join(
    cases_biweekly |> transmute(biweek, Cases = cases),
    by = 'biweek'
  )

corr_spearman <- biweekly_data |> 
  select(-biweek) |> 
  corrr::correlate(use = 'pairwise.complete.obs',
                   method = 'spearman')

# to get pvalues...
corrtest <- 
  biweekly_data |> 
  select(-biweek) |> 
  psych::corr.test(
    use = 'pairwise',
    method = 'spearman',
    adjust = 'none' # See p.adjust for "holm" > "bonferroni").
  )

# spearman r and p-values; upper triangle has adj. correlations
corr_r <- corrtest$r |> round(3)
corr_p <- corrtest$p |> round(3)

# remove lower tri with raw probabilities
corr_r[lower.tri(corr_r, diag = T)] <- NA
corr_p[lower.tri(corr_p, diag = T)] <- NA

corr_tbl_r <- kableExtra::kable(
  x = corr_r,
  format = 'pipe',
  caption = 'Campus-wide, biweekly metrics: Spearman r'
)
corr_tbl_p <- kableExtra::kable(
  x = corr_p,
  format = 'pipe',
  caption = 'Campus-wide, biweekly metrics: p-value on Spearman r'
)

corr_ci <- corrtest |> 
  pluck('ci') |>
  as_tibble() |> 
  mutate(terms = rownames(corrtest$ci), .before = 1L)

rm(corr_r, corr_p, corrtest)
```

</details>

<details>
<summary>Expand for plotting details</summary>

```{r figures, echo=TRUE}
# fig1 -----
fig1_timeline <- 
  event_dates |> 
  ggplot(aes(x = start, xend = end, y = event, yend = event)) +
  geom_segment(linewidth = 4, lineend = 'round',
               color = theme_color,
               alpha = 0.3) +
  geom_segment(linewidth = 1, lineend = 'butt',
             color = theme_color,
             alpha = 0.9) +
  geom_point(aes(x = end), color = theme_color, size = 1.6) +
  geom_point(color = theme_color, size = 1.6) +
  scale_x_study_dates() +
  labs(title = 'Timeline') +
  theme_no_x_labels() +
  theme_remove_grid()

fig1_cases <-
  cases_biweekly |>
  # adjust bar widths manually
  mutate(biweek = if_else(wday(biweek) == 5, biweek + 0.75, biweek - 0.75)) |>
  select(biweek, cases_student, cases_staff) |>
  pivot_longer(-biweek) |>
  mutate(name = str_remove(name, 'cases_') |> str_to_title()) |>
  ggplot() +
  geom_vline(
    data = tibble(x = as_date('2022-02-03')),
    aes(xintercept = x),
    lty = 2,
    color = 'gray'
  ) +
  geom_text(
    data = tibble(
      x = as_date('2022-02-05'),
      y = 18,
      label = 'End of data'
    ),
    aes(x, y, label = label),
    size = 2,
    hjust = 0
  ) +
  geom_col(
    aes(biweek, value, fill = name),
    size = 0.15,
    width = 3.5,
    color = 'gray80',
    position = position_stack()
  ) +
  labs(fill = NULL, title = 'UOttawa COVID-19 Cases') +
  scale_fill_carto_d() +
  scale_x_study_dates() +
  theme_no_x_labels() +
  theme_remove_grid() +
  theme(
    legend.position = c(0.9, 0.4),
    legend.key.size = unit(2, 'mm'),
    legend.background = element_rect(color = 'grey80')
  )

fig1_swabs <- 
  swabs_biweekly |> 
  ggplot(aes(biweek, positivity)) +
  geom_smooth(
    linewidth = 0.5, 
    alpha = 0.2, 
    fill = theme_color, 
    span = 0.4,
    method = 'loess',
    formula = 'y ~ x', 
    na.rm = T) +
  geom_point(color = theme_color, alpha = 0.85, shape = 1, size = 1) +
  scale_x_study_dates() +
  labs(title = 'Swab Positivity (%)') +
  theme_no_x_labels() +
  theme_remove_grid()
  
fig1_co2 <-
  ggplot(swabs_biweekly, aes(biweek, co2_mean)) +
  geom_point(color = theme_color, shape = 1, alpha = 0.85, size = 1) +
  geom_smooth(fill = theme_color, alpha = 0.2, size = 0.5) +
  labs(x = NULL, y = NULL, title = 'CO~2~ (ppm)') +
  scale_x_study_dates() +
  theme_no_x_labels() +
  theme_remove_grid()
  
fig1_wifi <-
  wifi_biweekly |> 
  ggplot(aes(biweek, mean)) +
  geom_smooth(span = 0.5, 
              size = 0.5,
              alpha = 0.2, 
              fill = theme_color) +
  geom_point(color = theme_color,  alpha = 0.85, shape = 1, size = 1) +
  labs(title = "Wi-Fi Connections") +
  scale_x_study_dates() +
  theme_no_x_labels() +
  theme_remove_grid()

fig1_uo_ww <-
  ggplot(uottawa_ww_biweekly,
         aes(biweek, signal)) +
  geom_smooth(span = 0.4, alpha = 0.2, size = 0.5,
              fill = theme_color, na.rm = T) +
  geom_point(
    # aes(size = n),
    alpha = 0.85, shape = 1, size = 1,
    color = theme_color,
    na.rm = T,
    show.legend = F) +
  labs(
    title = 'University Waste-Water',
  ) +
  scale_x_study_dates() +
  scale_size(range = c(1, 3)) +
  theme_no_x_labels() +
  theme_remove_grid()

fig1_ott_ww <- 
  ggplot(regional_ww_daily) + 
  geom_smooth(aes(sample_date, mean), 
              na.rm = T,
              method = 'loess', formula = 'y ~ x',
              span = 0.2, size = 0.5, alpha = 0.2,
              fill = theme_color) +
  geom_point(data = regional_ww_biweekly,
             aes(biweek, mean),
             na.rm = T,
             alpha = 0.85, shape = 1, size = 1,
             color = theme_color) +
  labs(x = 'Date',
       title = 'Regional Waste-Water',
       ) +
  scale_x_study_dates() +
  theme_remove_grid() +
  theme(axis.title.x = ggtext::element_markdown(lineheight = 1),
        axis.text.x = ggtext::element_markdown(size = 6.7)) 

figure_1 <- wrap_plots(
    fig1_timeline, fig1_cases, fig1_swabs, fig1_co2, 
    fig1_wifi, fig1_uo_ww, fig1_ott_ww
  ) +
  plot_layout(ncol = 1, nrow = 7, heights = c(1, rep(1.5, 6)))

rm(fig1_timeline, fig1_swabs, fig1_co2, fig1_wifi, fig1_uo_ww, fig1_ott_ww, fig1_cases)



## Figure 2: Corr heatmap -----------------------------------

figure_2 <- corrr::autoplot(corr_spearman, triangular = "lower",
                             barheight = 10) +
  geom_text(aes(label = r |> round(2)), size = 3.5)
  

## Figure 3: Per-building res ---------------------------------

#### thematic elements and axis scales -----

# limits are based on the min/max of data to keep scales consistent across all plots of a given variable
scale_y_co2 <- function() scale_y_continuous(limits = c(400, 950))
scale_y_copies <- function() scale_y_log10(limits = c(1, 300))
scale_y_cases <- function() {
  scale_y_continuous(breaks = seq(0, 5, 2), limits = c(0,5))
}
scale_y_wifi <- function() scale_y_continuous(limits = c(0, 1000))

scale_color_traffic <- function(){
  scale_color_carto_d(
    palette = 2,
    breaks = c('High', 'Low'),
    labels = c('High', 'Low'),
    drop = F
  )
}

theme_axes_lb <- function(){
  theme_remove_grid() +
  theme(
    panel.border = element_blank(),
    axis.line = element_line(colour = "gray30", linewidth = rel(0.5)),
    plot.margin = margin(2,2,2,2, 'mm')
  )
}

remove_y_axis_if_not_leftmost <- function(p, site){
  if (site %in% c('DMS', 'MRT')) return(p)
  p + theme_no_y_labels()
}

#### individual plot functions -----
plot_site_cases <- function(data, site){
  p <- data |>
    ggplot(aes(biweek, cases, 
               fill = factor(role, levels = c('Employee', 'Student')))
           ) +
    geom_col(linewidth = .1, width = 2.5, color = 'grey70') +
    geom_vline(data = tibble(xint = as_date('2022-02-02')),
               aes(xintercept = xint), lty = 2, alpha = 0.35) +
    scale_x_study_dates() +
    scale_y_cases() +
    scale_fill_carto_d(
      palette = 1,  
      breaks = c('Employee', 'Student'),
      labels = c('Employee', 'Student'),
      drop = FALSE,
      name = 'Role') +
    labs(fill = 'Cases', y = 'Cases', x = NULL, title = site) +
    theme_no_x_labels() +
    theme_axes_lb() +
    theme(axis.title.y = element_markdown()) 
  remove_y_axis_if_not_leftmost(p, site)
}

# swabs
plot_site_swabs <- function(data, site){
  data_early <- data |> filter(date_swab < '2022-01-01')
  data_late <- data |> filter(date_swab > '2022-01-01')
  
  p <- data_early |>
    ggplot(aes(date_swab, copies_plus_one, color = traffic)) +
    geom_path(alpha = 0.8, linewidth = 0.6) +
    geom_point(size = 0.6, alpha = 0.8) +
    geom_path(data = data_late, alpha = 0.8, linewidth = 0.6) +
    geom_point(data = data_late, size = 0.6, alpha = 0.8) +
    scale_x_study_dates() +
    scale_y_copies() +
    scale_color_traffic() +
    labs(x = NULL, y = 'Copies + 1', color = 'Traffic Level') +
    theme_no_x_labels() +
    theme_axes_lb() +
    theme(axis.title.y = element_markdown(),
          legend.text = element_markdown())
  remove_y_axis_if_not_leftmost(p, site)
}

# co2
plot_site_co2 <- function(data, site){
  data_early <- data |> filter(date_swab < '2022-01-01')
  data_late <- data |> filter(date_swab > '2022-01-01')
  
  p <- data_early |> 
    ggplot(aes(date_swab, co2, color = traffic)) +
    geom_path(alpha = 0.8) +
    geom_point(alpha = 0.8, size = 0.6) +
    geom_path(data = data_late, alpha = 0.8) +
    geom_point(data = data_late, alpha = 0.8, size = 0.6) +
    scale_x_study_dates() +
    scale_y_co2() +
    scale_color_traffic() +
    labs(y = 'CO~2~', x = NULL) +
    theme_no_x_labels() +
    theme_axes_lb() +
    theme(axis.title.y = element_markdown(), legend.position = 'none')
  remove_y_axis_if_not_leftmost(p, site)
}

# wifi
plot_site_wifi <- function(data, site){
  if (is.null(data)) return(patchwork::plot_spacer())
  
  p <- data |> 
    ggplot(aes(date, clients)) +
    geom_path(alpha = 0.5) +
    scale_x_study_dates() +
    scale_y_wifi() +
    labs(y = 'Wifi', x = NULL) +
    theme_axes_lb() +
    theme(
      axis.title.y = element_markdown(),
      axis.text.x = element_markdown(size = 6))
  remove_y_axis_if_not_leftmost(p, site)
}

## Plot datasets ====
fig3_cases <- 
  cases |> 
  unnest(associated_sites) |> 
  pivot_longer(UC:TBT, names_to = 'site') |> 
  mutate(site = if_else(site == 'UC', '90U', site)) |> 
  filter(value, 
         case_date > min(swabs_tidy$date_swab - 5),
         case_date < max(swabs_tidy$date_swab + 5),
         ) |> 
  group_by(site, biweek, role) |> 
  summarise(cases = n(),
            dates = list(case_date),
            .groups = 'drop') |>
  group_nest(site, .key = "cases")

fig3_swabs <- 
  swabs_tidy |> 
  mutate(
    site = str_remove(site, 'UO_'),
    traffic = get_traffic_level(location),
    copies = convert_cq_to_copies(pcr_ct),
    copies_plus_one = if_else(is.na(copies), 1, copies + 1),
    ) |> 
  select(site, traffic, date_swab, copies_plus_one) |>
  group_nest(site, .key = "swabs")

fig3_wifi <- 
  wifi |> 
  mutate(site = str_remove(site, 'UO_')) |> 
  select(date, site, clients) |>
  group_nest(site, .key = "wifi")

fig3_co2 <- 
   swabs_tidy |> 
  transmute(
    site = str_remove(site, 'UO_'),
    traffic = get_traffic_level(location),
    date_swab,
    co2,
    ) |> 
  group_nest(site, .key = "co2") |> 
  print()


#### data handling and making building panels -----

# combine nested df
fig3_df_nest <- fig3_cases |> 
  left_join(fig3_swabs, by = join_by(site)) |> 
  left_join(fig3_wifi, by = join_by(site)) |> 
  left_join(fig3_co2, by = join_by(site))

# map data to plots, combine into panel for each site
fig3_plts <- fig3_df_nest |>
  arrange(site == '90U') |> 
  transmute(
    site,
    plt_cases = map2(cases, site, plot_site_cases),
    plt_swabs = map2(swabs, site, plot_site_swabs),
    plt_co2 = map2(co2, site, plot_site_co2),
    plt_wifi = map2(wifi, site, plot_site_wifi),
  ) |> 
  mutate(panel = pmap(
    .l = list(site, plt_cases, plt_swabs, plt_co2, plt_wifi),
    .f = function(site,plt_cases, plt_swabs, plt_co2, plt_wifi) {
      p <- 
        patchwork::wrap_plots(
          plt_cases, plt_swabs, plt_co2, plt_wifi, ncol = 1
        ) 
      
      if (site %in% c('90U', 'MNT')) return(p)
      p + patchwork::plot_annotation(theme = theme(
        axis.title.y = element_blank(),
        axis.text.y = element_blank()
      ))
    }))

#### Combine all 6 sites' plots into 2*3 panel -----

figure_3 <- 
  patchwork::wrap_plots(
    fig3_plts$panel, 
    tag_level = 'keep',
    guides = 'collect'
  )


``` 

</details>

----

## Results

### Abstract Results

```{r abstract, results='asis'}
cat(rs_abstract)
```

---

<!-- ### Written Results -->

<!-- ```{r written results, results='asis'} -->
<!-- # cat(rs_results) -->
<!-- ``` -->

<!-- ---- -->

### Tables


#### Summary table

```{r}
# stack summaries
swab_summary |> 
  mutate(site = 'Overall') |> 
  bind_rows(swab_summary_sites) |> 
  relocate(site) |> 
  set_names(c('Site', 'N', 'PCR-Positive', '%')) |> 
  kableExtra::kable(
    format = 'pipe',
    caption = 'Building Summary'
  )
```


#### Spearman correlation between campus-wide measures

```{r}
kableExtra::kable(
  x = corr_ci, 
  format = 'pipe',
  caption = 'Spearman r for biweekly, campus-wide measures'
  )
```



---


### Figures

```{r fig.height=6, fig.width=5.25, fig.align='left'}
(figure_1) 
```

**Figure 1**: Time-series of (top to bottom) important events at the university, proportion of PCR-positive swabs, biweekly mean ambient CO~2~ (across collection sites), biweekly mean peak number of Wi-Fi connections (across 5 buildings), biweekly mean waste-water signal (across up to 6 collection sites; relative to PPMoV), biweekly mean regional waste-water detection (relative to PPMoV). Points show biweekly means. Trend lines show the LOESS fit.

----

<br>

```{r ts correlation, fig.height=4, fig.width=4, fig.align='left'}
figure_2
rm(corr_spearman)
```

**Figure 2**. Spearman correlation between biweekly campus-wide variables: self-reported cases, floor swab positivity (Swab PCR), mean CO2, mean daily peak Wi-Fi connections, aggregate waste-water signal at the university (University WW) and regional level (Ottawa WW), 


----



--------------------------------------------------------


```{r fig.width=10, fig.height=6}
figure_3
```
**Figure 3.** Campus COVID-19 cases, swab results, CO2 concentrations (during swab collection), and daily peak Wi-Fi connections by building. Cases plots show counts of self-reported  for students and employees separately; several cases are associated with multiple buildings. Case data collection was abandoned in early February 2022 (dashed line). Swabs and CO plots show results at two locations within each building, with one sample collected in a high-traffic area and the other in a low-traffic area. Swab PCR results are expressed as the number of SARS-CoV-2 RNA copies plus one, on a log-scale. Points represent the result for a single swab. Wi-fi plots show the peak daily number of simultaneous connections per building. No Wi-Fi data was available for the 90U building. 

 <!-- (Add shading for breaks, building closures). -->


```{r}
## save figures ----
# figs 1 & 3 don't work as a pdf for some reason
ggsave('fig/figure_1.png', figure_1, device = 'png', width = 7, height = 7) 
ggsave('fig/figure_2.pdf', figure_2, device = 'pdf', height = 4, width = 4)
ggsave('fig/figure_2.png', figure_2, device = 'png', height = 4, width = 4)
ggsave('fig/figure_3.png', figure_3, device = 'png', height = 8, width = 12)

```

------------------------------------------------------
  
<br>
  
```{r}
figure_sites_map
```

**Map: Where are these collection sites?** Pink markers represent waste-water collection sites; purple markers represent buildings where surface testing was performed. Only one building, Tabaret Hall, had both surface and waste-water testing. The locations of two waste-water sampling sites could not be ascertained and are not displayed on the map.

**It remains unclear** which buildings the 'South Sampling Site' and 'North Sampling Site' catchment areas are related to (not mapped).

--------------------------------------------

<br>

## Models

- **Hypothesis testing**. *H~A~*: swab-PCR positivity, CO~2~, wifi, and WW signals are  predictive of number of cases. *H~0~*: no relationship.
- **Method**: Backwards selection of significant predictors, ANOVA  
- **Formula**: `cases ~ swab_pcr + co2 + wifi + {university_ww} + regional_ww + (all interactions...)`. 
- **Issue**: only 10 / 50 biweekly obs are complete cases
   - `cases` counted only until Feb 2022 (~40% miss).  
   - `university_ww` from Dec 2021 (~40% miss).  
   - `wifi` occasional missing data. 
- **Question**: do we want standardized effects or simple effects? (I standardized because we are mostly interested in comparing different predictors within a model, as opposed to across different data sets.) 
- **Question**: do we want to consider interactions?  (I did and didn't)
   
<br>

```{r fig.height=5, fig.width=5, fig.align='center'}
# n=50 biweekly obs; uni_ww missing first half of period...
df_mod <- biweekly_data |> 
  janitor::clean_names() |> 
  mutate(across(-c(cases, biweek), scale)) # rescale

# only 10 complete observations, oof
df_mod_complete_obs <- df_mod |> na.omit()

# missing data
naniar::vis_miss(df_mod) +
  labs(subtitle = 'Aggregated TS Missingness')
```

<br>

--------------------------------------------

#### Case count distribution

**Cases counts** seem to follow a negative binomial distribution more than Poisson.

```{r fig.height=5, fig.width=5}
# simulated curves
add_poisson_density_curve <- function(p) {
  p + geom_density(
    data = tibble(
      x = rpois(n = sum(!is.na(df_mod$cases)),
                lambda = mean(df_mod$cases, na.rm = T))
      ), 
    aes(x), color = 'blue', alpha = 0.25, size = 0.03
  )
}
add_nb_density_curve <- function(p) {
  p +     
    geom_density(
      data = tibble(x = rnbinom(
        size = 1,
        n = sum(!is.na(df_mod$cases)),
        mu = mean(df_mod$cases, na.rm = T),
      )), 
      aes(x), color = 'blue', alpha = 0.25, size = 0.03
    )
}
simulation_plot <- function(plt_fn, title){
  p <- df_mod |> ggplot()
  walk(seq(100), ~ {p <<- plt_fn(p)})
  p + geom_density(aes(cases)) + labs(x = 'Cases', subtitle = title)
}

# Poisson dist
p1 <- simulation_plot(
  add_poisson_density_curve,
  'Cases vs. Simulated poisson distribution'
  )

# NB dist
p2 <- simulation_plot(
  add_nb_density_curve,
  'Cases vs. Simulated negative binomial distribution'
  )

(p1 / p2)

rm(p1, p2, add_poisson_density_curve, add_nb_density_curve)
```

<br>

--------------------------------------------

### Poisson Regression, Backwards Selection

<br>

#### Poisson Model with all interactions

Backwards selection arrived at the model with `swab_pcr`, `wifi`, `ottawa_ww`, and the interaction between `ottawa_ww` & `swab_pcr`.  

```{r}
# 30 complete obs when excluding university ww
df_mod_no_uniww <- df_mod |> select(-university_ww) |> na.omit()

# poisson model for cases with interactions
full_model <- glm(
  cases ~ . ^ 2,
  data = df_mod_no_uniww |> select(-biweek), 
  family = 'poisson')

# do backwards selection
backwards_selected <- step(full_model, trace = F)
stopifnot(backwards_selected$converged)

# standardized coefs and 95% CI
backwards_selected |> 
  broom::tidy(conf.int = T, conf.level = 0.95) |> 
  mutate(across(where(is.double), ~round(., 4))) |> 
  kableExtra::kable(
    format = 'pipe',
    caption = "Poisson model with interactions: Coefficients & 95% CI"
    )

# use type III SS tests since swab*ww interaction is significant
car::Anova(backwards_selected, type = 3) |> 
  rownames_to_column(var = 'Term') |> 
  as_tibble() |> 
    kableExtra::kable(
    format = 'pipe',
    caption = "Type III ANOVA"
    )

# test model assumption of equidispersion
# p > 0.05 means no significant overdispersion
AER::dispersiontest(backwards_selected)
```

<details>
<summary>***Backward selection model summary and diagnostics***</summary>
```{r}
summary(backwards_selected)
plot(backwards_selected)
hist(backwards_selected$residuals)

rm(full_model)
```
</details>

<br>

--------------------------------------------

#### Poisson Main Effects Only

Poisson model chosen by backwards selection with no interactions identifies swab positivity and Ottawa WW signal as significant predictors.

```{r}
# backwards select with n=30, no university ww
me_model <- glm(
  cases ~ swab_pcr + co2 + wi_fi + ottawa_ww,
  data = df_mod_no_uniww, 
  family = 'poisson')

backwards_selected_me <- step(me_model, trace = F)
stopifnot(backwards_selected_me$converged)

# nearly significant overdispersion when only main effects included
AER::dispersiontest(backwards_selected_me)

# regression coefs and 95%CI
backwards_selected_me |> 
  broom::tidy(conf.int = T, conf.level = 0.95) |> 
  mutate(across(where(is.double), ~round(., 3))) |> 
  kableExtra::kable(
    format = 'pipe',
    caption = "Main Effects Only: Coefficients & 95% CI"
    )

# type II SS test
car::Anova(backwards_selected_me, type = 2) |> 
  rownames_to_column(var = 'Term') |> 
  as_tibble() |> 
  kableExtra::kable(
    format = 'pipe',
    caption = "Type II ANOVA"
  )

```

<details>
<summary>***Backward selection model summary and diagnostics***</summary>
```{r}
summary(backwards_selected_me)
plot(backwards_selected_me)
hist(backwards_selected_me$residuals)
rm(me_model)
```
</details>


<br>

--------------------------------------------

### NB model, backwards selection

NB model with main effects also identifies `cases ~ swab + ottawa_ww` as optimal model (as with Poisson model). NB fit didn't converge when all interactions are included in the model.

<br>

#### Main effects, no interactions

```{r}
# negative binomial regression; main effects only
full_model_nb <- MASS::glm.nb(
  cases ~ ., 
  data = df_mod_no_uniww |> select(-biweek),
  control = glm.control(maxit = 5000)
  )

backwards_selected_nb <- step(full_model_nb, trace = F)
stopifnot(backwards_selected_nb$converged)

backwards_selected_nb |> 
  broom::tidy(conf.int = T, conf.level = 0.95) |> 
  mutate(across(where(is.double), ~round(., 4))) |> 
  rownames_to_column(var = 'Term') |> 
  kableExtra::kable(
    format = 'pipe',
    caption = "NB model -no interactions: Coefficients & 95% CI"
    )
```

<details>
  <summary>***Model summary and diagnostics***</summary>
```{r}
# not the worst fit, could be much better
summary(backwards_selected_nb)
plot(backwards_selected_nb)
hist(backwards_selected_nb$residuals)
rm(full_model_nb)
```

</details>


The negative binomial model (with extra parameter $\theta$) does not significantly improve the log likelihood compared to the poisson model (p > 0.05).
```{r}
library(lmtest)
# poisson vs. nb models with main effects only
lmtest::lrtest(backwards_selected_me, backwards_selected_nb)
```



<details>
  <summary>**NB fit fails with interactions**</summary>

```{r echo=T}
# swab pcr and 
full_model_nb_int <- MASS::glm.nb(
  cases ~ .*., 
  data = df_mod_no_uniww |> select(-biweek) |> na.omit(),
  control = glm.control(maxit = 15000)
  )
# do selection
try((backwards_selected_nb_int <- step(full_model_nb_int)))
rm(full_model_nb_int)

```
</details>

<br>

--------------------------------------------

### Model with university WW

If we consider the ten biweekly observations where we have linked cases, swabs, and university WW data,

```{r}
# backwards select with university ww (n=10 biweekly obs)
full_model_w_uni_ww <- glm(
  cases ~ .,
  data = df_mod_complete_obs |> select(-biweek), 
  family = 'poisson')

# do backwards selection
backwards_selected_ww <- step(full_model_w_uni_ww, trace = F)
stopifnot(backwards_selected_ww$converged)


# standardized coefs and 95% CI
backwards_selected_ww |> 
  broom::tidy(conf.int = T, conf.level = 0.95) |> 
  mutate(across(where(is.double), ~round(., 4))) |> 
  kableExtra::kable(
    format = 'pipe',
    caption = "Poisson model with interactions: Coefficients & 95% CI"
    )

# use type II SS test
car::Anova(backwards_selected_ww, type = 2) |> 
  rownames_to_column(var = 'Term') |> 
  as_tibble() |> 
    kableExtra::kable(
    format = 'pipe',
    caption = "Type II ANOVA"
    )

# test model assumption of equidispersion
# p > 0.05 means no significant overdispersion
AER::dispersiontest(backwards_selected_ww)
```


<details>
  <summary>***Model summary and diagnostics***</summary>
```{r}
summary(backwards_selected_ww)
plot(backwards_selected_ww)
hist(backwards_selected_ww$residuals)
rm(full_model_w_uni_ww)
```

</details>




--------------------------------------------

### GLMM: High v Low Traffic

```{r}
swab_summary_location_traffic |> 
  set_names(c('Traffic level', 'N', 'PCR-Positive', '%')) |> 
  kableExtra::kable(
    format = 'pipe',
    caption = 'Positivity by sample location traffic-level'
  )
```


To evaluate whether high traffic locations are different from low traffic sites, in terms of swab-PCR positivity, we fit a mixed model with random intercepts for different sites. There appears to be little difference (15% vs 12% overall) and this is confirmed by our mixed model with traffic level as a fixed effect (OR 0.71; 95% CI 0.82-2.2). 


```{r}
traffic_swabs <- 
  swabs_tidy |> 
  mutate(
    traffic = get_traffic_level(location) |> 
      str_remove('\\straffic') |> 
      factor(levels = c('low', 'high'))
    ) |> 
  select(site, traffic, pcr_positive)

mixed_model <- 
  lme4::glmer(pcr_positive ~ traffic + (1 | site), data = traffic_swabs,
            family = 'binomial')

mixed_model |> 
  broom.mixed::tidy(conf.int = T, exponentiate = T) |> 
  mutate(across(where(is.double), ~round(., 4))) |> 
  kableExtra::kable(
    format = 'pipe',
    caption = 'Traffic level - Random intercepts model'
  )

```


--------------------------------------------


### Model: Cases w/in buildings

```{r}
study_period_weeks <- crossing(
  date = as_date(as_date('2021-09-21'):as_date('2022-01-31')),
  site = c('90U', 'DMS', 'LPR', 'MNT', 'MRT', 'TBT')
  ) |> 
  mutate(
    week = get_date_week(date),
    redate = week_to_date(week)) |> 
  distinct(site, week)

# group swabs by week & site
per_bldg_swabs <- 
  swabs_tidy |> 
  filter(date_swab < '2022-02-02') |> 
  mutate(
    site = str_remove(site, 'UO_'),
    week = get_date_week(date_swab) |> as.character(),
  ) |> 
  group_by(site, week) |> 
  summarise(
    positivity = sum(pcr_positive == 'Positive') / n(),
    .groups = 'drop')

# 72 cases associated with study bldgs (at least one)
per_bldg_cases <- 
  cases |> 
  filter(case_date > '2021-09-20') |> 
  unnest(associated_sites) |> 
  pivot_longer(UC:TBT, names_to = 'site') |> 
  filter(value) |> 
  select(-value, -biweek, -case) |> 
  mutate(
    site = str_replace(site, 'UC', '90U'),
    week = get_date_week(case_date)) |> 
  group_by(week, site) |> 
  summarise(cases = n(), .groups = 'drop') 

per_bldg_df <- per_bldg_swabs |> 
  left_join(per_bldg_cases, by = join_by(site, week)) |> 
  mutate(cases = if_else(is.na(cases), 0, cases))

rm(study_period_weeks, per_bldg_cases, per_bldg_swabs)
```


**Poisson GLM w random intercepts: Cases ~ positivity + (1|site)**

```{r}
per_bldg_model <- lme4::glmer(
  cases ~ positivity + (1|site),
  data = per_bldg_df,
  family = 'poisson'
)
broom.mixed::glance(per_bldg_model)
# Wald CIs
broom.mixed::tidy(per_bldg_model, conf.int = T, conf.level = 0.95)

```


```{r}
points_df <- crossing(
  positivity = seq(0, 1, 0.1),
  site = unique(per_bldg_df$site) 
)

points_df |> 
  mutate(   
    pred = predict(
      type = 'response',
      object = per_bldg_model, 
      newdata = points_df
    )) |> 
  ggplot(aes(positivity, pred, color = site)) +
  geom_smooth(se = F, linewidth = 0.25, alpha = 1) + 
  geom_jitter(data = per_bldg_df,
              alpha = 0.75, shape = 1, size = 0.85,
              width = 0.01, height = 0.01,
              aes(positivity, cases)) +
  labs(x = 'Swab positivity', y = 'Cases', color = 'Site')
```


```{r}
broom.mixed::tidy(per_bldg_model, 'ran_vals', 
                  conf.int = T, conf.level = 0.95) |> 
  ggplot(aes(y = level,
             x = estimate, xmin = conf.low, xmax = conf.high)) +
  geom_pointrange()
```


--------------------------------------------

## Case Data Imputation... 

Multiple imputation (mice) has a key assumption that data is MAR. But data is (arguably) not MAR, but because of some dynamics not measured in the data (like the university giving up hope on recording the data in 2022). 

Using `mice` means making the analysis *much* more complicated... Generally, you need to model each imputed dataset and then analyze the pooled results. That gets a lot more complicated when needing to first aggregate the data, join to a bunch of other data, do various modelling tasks.... all nested within each iteration.

Instead, I used a simpler ad-hoc method:  

  - if positive-test-result date is available, use that, else:  
  - use -5 days from end-of-isolation date, or if not available,  
  - use +3 days from symptoms-began date  
  
It's not really the most sound approach, but the data is a bit questionable to begin with anyway...


<!-- ```{r} -->
<!-- read_csv(here('data/uo_cases_to_date.csv')) |>  -->
<!--   select(positive_result, symptoms_began, tested, isolation_end) |>  -->
<!--   pivot_longer(-positive_result) |>  -->
<!--   ggplot(aes(value, positive_result)) + -->
<!--   geom_smooth(method = 'lm', alpha = 0.5) + -->
<!--   geom_point(size = 0.5) + -->
<!--   facet_grid(~name) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- read_csv(here('data/uo_cases_to_date.csv')) |>  -->
<!--   filter(is.na(positive_result)) |>  -->
<!--   select(symptoms_began, tested, isolation_end) |>  -->

<!--   pivot_longer(everything()) |>  -->
<!--   ggplot(aes(value)) + -->
<!--   geom_histogram() + -->
<!--   facet_grid(~name) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- read_csv(here('data/uo_cases_to_date.csv')) |>  -->
<!--   # 28 have no result date -->
<!--   filter(is.na(positive_result)) |>  -->
<!--   # 3 of 28 are also missing isolation-end-date -->
<!--   filter(is.na(isolation_end)) |>  -->
<!--   # 0 are missing result, isolation, and symptoms! -->
<!--   filter(is.na(symptoms_began)) |> -->
<!--   glimpse() -->

<!-- read_csv(here('data/uo_cases_to_date.csv')) |>  -->
<!--   filter(is.na(positive_result), is.na(tested))  -->
<!-- ``` -->

--------------------------------------------



<details>

<summary>R Software details</summary>

- Statistical analysis was performed using  R version 4.2.2 (2022-10-31; cite Ihaka & Gentleman).

- Generalized linear models were created using the `glm` function. Mixed models were created using the `glmer` function from the package `lme4` (Bates et al.).

- We fit poisson and negative binomial regression models with the number of campus-wide cases as the outcome; we applied backwards selection to evaluate predictors of campus-wide cases, including the aggregated results (means) of surface swabbing, waste-water testing, CO2 monitors, and Wi-Fi user counts.

- We used mixed effects logistic regression to model the presence of SARS-CoV-2 infected individuals as a binary outcome (ie. positive event: one or more cases occurring during a week) in a university building, using surface swab PCR results as predictor with random intercepts for buildings.

- Graphics were created with ggplot2 (v3.4.1)

- Find our analysis code at our public github repository: https://github.com/CUBE-Ontario/UOttawa-Analysis


```{r}
sessionInfo()
```

</details>