---
title: "CUBE - UOttawa"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_chunks: hide
    toc: true
    toc_float: true
    theme: spacelab
---

```{r setup, echo=F}
knitr::opts_chunk$set(echo = F, message = F, warning = F)

```


```{r}
library(tidyverse)
library(here)
library(ggiraph)
library(patchwork)
ggplot2::theme_set(theme_bw())
source(here('R/report_plots.R'))


# creates yyyy-ww label for grouping data
get_date_week <- function(x){
  y <- lubridate::year(x)
  w <- lubridate::week(x) |> str_pad(2, 'left', 0)
  str_glue("{y}-{w}")
}

# lookup table for date
week_midpoint_date_lookup <- function(start = '2022-01-01', end = '2025-01-01'){
  ends = lubridate::as_date(c(start, end))
  tibble(
    date = seq(ends[1], ends[2], by = 1),
    date_week = lubridate::week(date) |> str_pad(2, 'left', '0'),
    date_year = lubridate::year(date),
    week = str_glue("{date_year}-{date_week}"),
  ) |> 
    group_by(week) |> 
    summarise(date = mean(date))
}
binom_ci <- function(x, n){
  Hmisc::binconf(x, n, method = 'wilson', alpha = 0.05) |> 
    as_tibble() |> 
    janitor::clean_names()
}
get_binom_ci <- function(data){
  data |> 
    summarise(
      x = sum(pcr_positive == 'Positive'),
      n = n(),
      binconf = map2(x, n, ~binom_ci(.x, .y))
    ) |> 
    unnest(binconf) |> 
    mutate(across(c(point_est, lower, upper), ~(.x*100) |> round(1)))
}
```


```{r}
# swab data
swabs <- read_rds(here('data/cube.rds')) |> 
  filter(str_detect(site, '^UO_')) 

# waste water data 
raw_ww <-
  readxl::read_excel(here::here('data/wastewater.xlsx')) |>
  janitor::clean_names()

ww_lookup <- tribble(
  ~site_abbr, ~site,
  'UO_AA',	'Annex Residence',
  'UO_FA',	'Faculty of Social Sciences',
  'UO_FT',	'Friel Residence',
  'UO_na',	'Northern sampling site',
  'UO_RGN',	'Roger Guindon Hall',
  'UO_ST',	'Southern sampling site',
  'UO_TBT',	'Tabaret Hall'
)

# OttawaU data from R. Delatolla
uottawa_ww <-
  raw_ww |>
  filter(!is.na(sample_date)) |>
  mutate(
    source =
      source_name |>
      str_remove('^Data - uOttawa - ') |>
      str_remove('.xlsx$'),
    site = case_when(
      source == 'uo_aa' ~ 'Annex Residence',
      source == 'uo_fa' ~ 'Faculty of Soc. Sci.',
      source == 'uo_ft' ~ 'Friel Residence',
      source == 'uo_na' ~ 'North sampling Site',
      source == 'uo_rgn' ~ 'Robert Guidon Hall',
      source == 'uo_st' ~ 'South Sampling Site',
      source == 'uo_tbt' ~ 'Tabaret Hall',
      TRUE ~ 'not matched'
    )
  ) |> 
  filter(
    sample_date > '2021-09-15', 
    sample_date < '2022-05-05',
    site != 'Robert Guidon Hall',
    !is.na(sample_date),
    !is.na(site),
    !is.na(viral_copies_per_copies_pep_avg)
    ) |> 
  transmute(site, sample_date, 
            signal = viral_copies_per_copies_pep_avg)

# ottawa wastewater data: summary
regional_wastewater <- 
  read_rds(here('data/ww_ottawa.rds')) |> 
  select(sample_date, starts_with('cov_')) |> 
  pivot_longer(contains('cov_')) |> 
  mutate(target = str_extract(name, 'cov_n.'),
         stat = str_extract(name, 'mean|sd'),
         ) |> 
  select(-name) |> 
  pivot_wider(names_from = stat, values_from = value) |> 
  mutate(.lower = mean - sd, .upper = mean + sd)   

wifi <- 
  read_rds(here('data/uo_wifi.rds')) |> 
  filter(date >= min(swabs$date_swab), date <= max(swabs$date_swab))

```

----

## Where

Are these buildings?

```{r}

swab_coords <- tribble(
  ~site, ~name, ~lat, ~lng,
  'MRT', 'Morrisette Hall (MRT)', 45.423239101483546, -75.68428970961207,
  'MNT', 'Montpetit', 45.42254171025894, -75.68265871464943,
  '90U', '90 University', 45.42242555707402, -75.6850144991752,
  'DMS', 'Desmarais Bldg.', 45.42387679340988, -75.68727075448753,
  'TBT', 'Tabaret Hall', 45.42450940162542, -75.68631900184072,
  'LPR',  'Louis Pasteur Bldg.', 45.421375668614466, -75.68026389993058,
)

ww_coords <- tribble(
  ~site, ~name, ~lat, ~lng,
  'aa', 'Annex Residence', 45.42678129026445, -75.68034405960745,
  'fa', 'Faculty of Soc. Sci.', 45.421624722628515, -75.68390386012699,
  'ft', 'Friel Residence', 45.43043440080566, -75.68290766533741,
  'na', 'North sampling Site', NA, NA,
  'st', 'South Sampling Site', NA, NA,
  'tbt', 'Tabaret Hall', 45.42450940162542, -75.68631900184072
)

leaflet::leaflet(swab_coords, width = 600, height = 330) |> 
  leaflet::addProviderTiles(leaflet::providers$CartoDB.Positron) |> 
  leaflet::setView(lng = -75.6843, lat = 45.4235, zoom = 15) |> 
  leaflet::addCircleMarkers(~lng, ~lat, label =  ~name, popup = ~name, 
                            radius = 2, color = '#440099') |> 
  leaflet::addLabelOnlyMarkers(
    ~lng, ~lat, label =  ~site, 
    labelOptions = leaflet::labelOptions(
      noHide = T, direction = 'top', textOnly = T, 
    )) |> 
  leaflet::addCircleMarkers(~lng, ~lat, label =  ~site, 
                            radius = 2, color = '#440099') |> 
  leaflet::addCircleMarkers(
    data = ww_coords, ~lng, ~lat, label =  ~name, popup =  ~name, 
    radius = 4, color = '#f96999')
  
```


For WW data, I am still unsure which buildings the 'South Sampling Site' and 'North Sampling Site' catchment areas are related to.

-----

## Waste Water - UO

Last time, I was confused about which buildings we had obtained waste-water data for. This week, I've removed the off-campus site (Robert Guidon Hall) and retained the other 6 on-campus sites (raw data below).

```{r fig.width=8, fig.height=7}
uottawa_ww |>
  ggplot(aes(sample_date, signal)) +
  geom_path(alpha = 0.6) +
  geom_point(alpha = 0.5, size = 0.5) +
  labs(x = 'Date', y = 'SARS-CoV-2 / PPMoV') +
  facet_wrap(~site, ncol = 2)

```

## Waste Water - Regional

```{r}
regional_wastewater |> 
  ggplot() +
  geom_pointrange(
    aes(sample_date, mean, color = target, fill = target,
        ymin = .lower, ymax = .upper),
    size = 0.25, fatten = 1, alpha = 0.25
    ) +
  geom_smooth(
    aes(sample_date, mean, color = target, fill = target),
    span = 0.2, alpha = 0.15
  ) +
  geom_rect(data = swabs |> summarise(xmin = min(date_swab),
                                      xmax = max(date_swab),
                                      ymin = -Inf, 
                                      ymax = Inf),
            aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
             lty = 2, alpha = 0.1) +
  geom_text(aes(x = min(swabs$date_swab) + 5, y = 0.0035, 
                label = 'UOttawa study'),
            color = 'grey50', hjust = 0) +
  scale_shape(solid = F) +
  labs(x = 'Date', y = 'SARS-CoV-2 Target / PPMoV ')
```



----

<br>

We probably want to summarise our swabs at the weekly (or biweekly?) level for comparison to the waste-water signal. This is what the weekly campus-wide time-series looks like compared with our weekly campus-wide swab positivity:


```{r fig.width = 6, fig.height=6}
ww_weekly_summary <- 
  uottawa_ww |> 
  mutate(week = get_date_week(sample_date)) |> 
  left_join(week_midpoint_date_lookup('2021-01-01', '2022-12-31'),
            by = 'week') |> 
  select(date, signal, site) |> 
  group_by(date) |> 
  summarise(signal_mean = mean(signal),
            signal_sd = sd(signal))

plt_ww_weekly <- ww_weekly_summary |> 
  ggplot(aes(date, 
             signal_mean, 
             # ymin = signal_mean - signal_sd,
             # ymax = signal_sd + signal_sd
             )) +
  geom_line() +
  geom_point() +
  geom_point(data = uottawa_ww,
             aes(as_date(sample_date), signal),
             size = 1, shape = 1, alpha = 0.5) +
  labs(x = 'Date', y = 'SARS-CoV-2 / PMMoV',
       title = 'Weekly campus-wide average SARS-CoV-2 waste-water detection')

swabs_weekly_summary <- swabs |> 
  mutate(week = get_date_week(date_swab)) |> 
  left_join(week_midpoint_date_lookup('2021-01-01', '2022-12-31'),
            by = 'week') |> 
  group_by(date) |> 
  summarise(
    x = sum(pcr_positive == 'Positive'),
    n = n(),
    positivity = list(binom_ci(x, n))
    ) |> 
  unnest_wider(positivity)

plt_swabs_weekly <- swabs_weekly_summary |> 
  ggplot(aes(x = date, y = point_est*100)) +
  geom_line() +
  geom_point() +
  scale_x_date() +
  scale_y_continuous(limits = c(0,100)) +
  labs(x = 'Date', y = 'Swab positivity (%)',
       title = 'Weekly campus-wide surface swab detection rate')

wrap_plots(plt_swabs_weekly, plt_ww_weekly, ncol = 1)
```

If we look at the relationship between swabs and waste-water on weeks where both results are available, we see that they have a positive correlation.

```{r fig.width = 5, fig.height=5}
campus_ww_swab_summary <- 
  swabs_weekly_summary |> 
  select(date, point_est) |> 
  left_join(ww_weekly_summary, by = 'date') |> 
  transmute(date, ww = signal_mean, swabs = point_est*100)

campus_ww_swab_summary |> 
  ggplot(aes(swabs, ww, color = date)) +
  geom_smooth(method = 'lm', color = 'gray', fill = 'gray') +
  geom_point() +
  scale_color_date() +
  coord_fixed(ratio = 18000) +
  labs(x = 'Swab positivity (%)', 
       y = 'Waste-water signal',
       color = "Sample\nWeek")
```


```{r}
cortest_swab_ww <- 
  lst(
    pearson = cor.test(
      campus_ww_swab_summary$ww,
      campus_ww_swab_summary$swabs,
      method = 'pearson'
    ),
    pearson_est = pearson$estimate |> round(2),
    pearson_lower = pearson$conf.int[1] |> round(2),
    pearson_upper = pearson$conf.int[2] |> round(2),
    
    spearman = cor.test(
      campus_ww_swab_summary$ww,
      campus_ww_swab_summary$swabs,
      method = 'spearman'
    ),
    spearman_est = spearman$estimate |> round(3),
    # spearman_lower = spearman$conf.int[1] |> round(2),
    # spearman_upper = spearman$conf.int[2] |> round(2),
    
    kendall = cor.test(
      campus_ww_swab_summary$ww,
      campus_ww_swab_summary$swabs,
      method = 'kendall'
  ),
    kendall_est = kendall$estimate |> round(2),
    # kendall_lower = kendall$conf.int[1] |> round(2),
    # kendall_upper = kendall$conf.int[2] |> round(2),
    
) |> 
  suppressWarnings() |> 
  glue_data(
    "The Pearson's *r* (linear correlation) is {pearson_est} ({pearson_lower}-{pearson_upper}).
    The Spearman's *r* (rank correlation) is {spearman_est}..
    The Kendall's *Tau*  is {kendall_est}."
  )

```

`r cortest_swab_ww`

Pearson's *r* is more affected by outliers. Spearman's *r* or Kendall's *Tau* are probably the more reliable measures of correlation between these variables.

<br>
<br>

(Data doesn't appear normally distributed, btw)

```{r fig.height=3, fig.width=7.5}
campus_ww_swab_summary |> 
  pivot_longer(cols = c(ww, swabs)) |> 
  ggplot(aes(value)) +
  geom_histogram(bins = 15, color = 'lightgray') +
  facet_wrap(~name, scales = 'free')
```

<br>
<br>

------------------------------------------------------------------------

## 📊 Summary

This plot shows the counts of positive (red) and negative (yellow) samples collected at each facility over time (x-axis). Samples that could not be tested are shown in navy. Only flocked swabs are shown. (Other sponge swabs were collected on 2022-04-28 were for comparing flocked and sponge swabs.)

```{r university bargraph}
plot_timeseries_summary(
  swabs |> 
    filter(
      swab_type %in% 'flock',
      !negative_control,
      !is.na(negative_control)
    ),
  height_svg = 3.3,
  width_svg = 8
  )
```


------------------------------------------------------------------------

## ⡯⡷ Dotplot

This plot shows the counts of positive and negative samples collected at each facility over time. 

```{r univeristy dotplot}
plot_timeseries_locations(
  swabs |> 
    filter(swab_type == 'flock'),
  height_svg = 6,
  width_svg = 8
  )
```

----------------------------------------------------------------------


## Model {.tabset}

This section contains results from modeling SARS-CoV-2 cases at UO using swab-PCR results as a predictor.

<br>


#### Specification

We created a random intercepts logistic regression model with the occurrence of cases (binary) as outcome and swab results for the previous week (the proportion of positive swabs) as predictor. The model has a random intercept for each site.

Our model formula is `cases ~ positives[lag 1 week] + (1|site)`.

The model is fit as follows:

```{r eval=FALSE, echo=TRUE}
swab_model <-
  blme::bglmer(
    cases_binary ~ detection_lag_1week + (1 | site),
    data =  uo_sites,
    family = binomial
  )
```

------------------------------------------------------------------------

```{r model}
swab_model <- read_rds(here('model/uo_swab_model.rds'))
```

<br>

#### Model response time-series

These plots show the swab results, cases, and predictions by the current model.\
<br>

```{r prediction figure live version, fig.height=15, fig.width=8}
# uo cases data - long table
cases_long <-
  read_rds(here('./data/latest_uo_cases.rds')) |>
  filter(!is.na(positive_result),
         positive_result > lubridate::as_date('2021-09-14')) |>
  select(case, role, confirmed_by_test, positive_result, c(`90U`:TBT)) |>
  mutate(
    `Campus Total` = TRUE,
    # 3 days transmissibility
    transmission_start = positive_result - days(3)
    ) |>
  pivot_longer(`90U`:`Campus Total`, names_to = 'site') |>
  filter(value) |>
  select(-value) |> 
  mutate(site = fct_relevel(site, "Campus Total", after = 0L))

uo_pred <- read_rds(here('model/uo_pred.rds'))

# create a 3-panel fig for each of 6 buildings.
map(
  # site names for filtering 
  .x = swabs$site |> unique() |> str_remove('UO_'),
  .f = ~plot_uo_pred_swab_case(
    swabs = swabs,
    cases = cases_long,
    preds = uo_pred,
    site_select = .x
  )
) |> wrap_plots(guides = 'collect', ncol = 1)

```

<br>

------------------------------------------------------------------------

#### Model summary

These tables show the model coefficients and statistics.

```{r model glance}
broom.mixed::glance(swab_model) |> 
  mutate_if(is.numeric, round, 2) |> 
  kableExtra::kable(caption = 'Model statistics',
                    format = "html") |> 
  kableExtra::kable_styling(bootstrap_options = "striped", 
                            full_width = F, position = "left")
```

------------------------------------------------------------------------

```{r model tidy}
broom.mixed::tidy(swab_model) |>
  mutate(p.value = as.character(p.value)) |> 
  mutate_if(is.numeric, round, 3) |> 
  mutate(p.value = as.numeric(p.value)) |> 
  kableExtra::kable(caption = 'Model parameters',     
                    format = "html") |> 
  kableExtra::kable_styling(bootstrap_options = "striped",
                            full_width = F, position = "left")
# tidy(swab_model, effects = 'ran_vals')
```

<br>

------------------------------------------------------------------------

```{r plot model intercepts}
sjPlot::tab_model(
  swab_model,
  show.re.var = TRUE,
  pred.labels = c("(Intercept)", "Swabs @ t-1week"),
  dv.labels = "Relation between UO cases (y/n) and proportion\n of positive swabs the previous week"
)
```

<br>

------------------------------------------------------------------------

#### Model response curves

This plot shows the modelling data as points (detection lagged 1 week and cases at a given site- y/n), as well as how the probability of future cases varies by the previous weeks detection level and site, according to our model (curves).

```{r response curves, fig.height=4.5}
read_rds(here('fig/plt_model_resp_curves.rds')) +
  labs(subtitle = 'Points show case/swab data and curves show model probability values')
```

<br>

------------------------------------------------------------------------

#### Random Intercepts for Sites

This plot shows the odds ratios for the intercepts of the model (an intercept for each site).

```{r model_sites}
# random effect plot: site intercepts; axes are flipped
sjPlot::plot_model(swab_model, type = 're') +
  theme_light() +
  labs(title = "Random effects", y = 'Odds Ratio')
```

<br>

------------------------------------------------------------------------

### Swabs and Cases

This panel shows the cases counts at UO over the course of our sampling period. The case data shown represents the days on which a positive test was reported (black rug lines) and the presumed start of transmissiblity for each case (red lines).

```{r cases and swabs, fig.height=6, fig.width=8, fig.align='center'}
# generated from file R/plot_UO_case_load.R
read_rds(here('fig/plt_cases_swabs.rds'))
```

<br>

------------------------------------------------------------------------

### Swabs, Wifi, and CO~2~

This panel shows linked data from swab results, CO~2~ readings, and wifi traffic (number of users \@ peak, daily) during our study period. Unfortunately, we do not have wifi data for 90U. <br>

```{r uo multivariate panel, fig.width=12.5, fig.asp=0.6}
uo_multivariate <- make_uo_panel(swabs, wifi)
write_rds(uo_multivariate, here('fig/uo_multivariate.rds'))

uo_multivariate & theme(text = element_text(size = 13))
rm(uo_multivariate)
```

<br>

------------------------------------------------------------------------

### Wifi Traffic

```{r wifi_ts, fig.height=6, fig.width=8}

plot_UO_wifi_ts(wifi, swabs)

```

This panel shows a time-series of the daily peak number of wifi users at UO facilities. Sampling days are highlighted in blue.\
<br>

### 