---
title: 'CUBE: UOttawa EDA'
author: "Jason Moggridge"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_fold: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
```


```{r}
library(tidyverse)
source('R/report_plots.R')

## functions ------

# creates yyyy-ww label for grouping data
get_date_week <- function(x){
  y <- lubridate::year(x)
  w <- lubridate::week(x) |> str_pad(2, 'left', 0)
  str_glue("{y}-{w}")
}

# get biweekly bin date for vector of dates
get_date_biweekly <- function(x){
  day_of_week <-  lubridate::wday(x)
  case_when(
    day_of_week %in% 1:3 ~ x + 3 - day_of_week,
    TRUE ~ x + 5 - day_of_week,
  )
}

# lookup table for date
week_midpoint_date_lookup <- function(start = '2021-01-01', end = '2023-01-01'){
  ends = lubridate::as_date(c(start, end))
  tibble(
    date = seq(ends[1], ends[2], by = 1),
    date_week = lubridate::week(date) |> str_pad(2, 'left', '0'),
    date_year = lubridate::year(date),
    week = str_glue("{date_year}-{date_week}"),
  ) |> 
    group_by(week) |> 
    summarise(date = mean(date))
}


binom_ci <- function(x, n){
  Hmisc::binconf(x, n, method = 'wilson', alpha = 0.05) |> 
    as_tibble() |> 
    janitor::clean_names()
}

get_binom_ci <- function(data){
  data |> 
    summarise(
      x = sum(pcr_positive == 'Positive'),
      n = n(),
      binconf = map2(x, n, ~binom_ci(.x, .y))
    ) |> 
    unnest(binconf) |> 
    mutate(across(c(point_est, lower, upper), ~(.x*100) |> round(1)))
}


## data ----------

swabs <- 
  read_rds(here('data/cube.rds')) |> 
  filter(str_detect(site, '^UO_')) 

# data from university of ottawa
wifi <- 
  read_rds(here('data/uo_wifi.rds')) |> 
  filter(date >= min(swabs$date_swab), 
         date <= max(swabs$date_swab))

# lookup table for waste water site names
lookup_ww <- tribble(
  ~site_abbr, ~site,
  'UO_AA',	'Annex Residence',
  'UO_FA',	'Faculty of Social Sciences',
  'UO_FT',	'Friel Residence',
  'UO_NA',	'Northern sampling site',
  'UO_RGN',	'Roger Guindon Hall',
  'UO_ST',	'Southern sampling site',
  'UO_TBT',	'Tabaret Hall'
)

# UOttawa waste water data from R. Delatolla
uottawa_ww <-
  readxl::read_excel(here::here('data/ww_university.xlsx')) |>
  janitor::clean_names() |> 
  mutate(sample_date = as_date(sample_date)) |> 
  filter(
    sample_date > '2021-09-15', 
    sample_date < '2022-05-05',
    !is.na(viral_copies_per_copies_pep_avg),
    site != 'UO_RGN'
    ) |> 
  left_join(lookup_ww, by = c('site' = 'site_abbr')) |> 
  mutate(
    source =
      source_name |>
      str_remove('^Data - uOttawa - ') |>
      str_remove('.xlsx$')
  ) |> 
  transmute(site,
            sample_date, 
            signal = viral_copies_per_copies_pep_avg)

# ottawa wastewater data: daily means
regional_ww <- 
  read_rds(here('data/ww_ottawa.rds')) |> 
  select(sample_date, starts_with('cov_')) |> 
  pivot_longer(contains('cov_')) |> 
  mutate(target = str_extract(name, 'cov_n.'),
         stat = str_extract(name, 'mean|sd'),
         ) |> 
  select(-name) |> 
  pivot_wider(names_from = stat, values_from = value) |> 
  mutate(.lower = mean - sd, .upper = mean + sd)   

# important dates for context
event_dates <- 
  tribble(
    ~event, ~start, ~end,
    'Reading Week',  '2021-10-24', '2021-10-30',
    'Holiday Break', '2021-12-22', '2022-01-04',
    'Closure',       '2022-01-04', '2022-01-31',
    'Closure',       '2022-02-16', '2022-02-21',
    'Reading Week',  '2022-02-20', '2022-02-26',
  ) |> 
  mutate(across(start:end, as_date))

# UOttawa cases data
cases <-
  read_rds(here('data/cases_rule_imputed.rds')) |> 
  select(case, role, case_date, UC:TBT) |>
  mutate(biweek = get_date_biweekly(case_date)) |> 
  nest(associated_sites = UC:TBT)
```

------------------------------------------------------------------------

## 📊 Summary

This plot shows the counts of positive (red) and negative (yellow) samples collected at each facility over time (x-axis). Samples that could not be tested are shown in navy. Only flocked swabs are shown. (Other sponge swabs were collected on 2022-04-28 were for comparing flocked and sponge swabs.)

```{r university bargraph}
plot_timeseries_summary(
  swabs |>
    filter(
      swab_type %in% 'flock',
      !negative_control,
      !is.na(negative_control)
    ),
  height_svg = 3.3,
  width_svg = 8
  )
```


------------------------------------------------------------------------

## ⡯⡷ Dotplot

This plot shows the counts of positive and negative samples collected at each facility over time.

```{r univeristy dotplot}
plot_timeseries_locations(
  swabs |>
    filter(swab_type == 'flock'),
  height_svg = 6,
  width_svg = 8
  )
```

----------------------------------------------------------------------


### Model {.tabset}

This section contains results from modeling SARS-CoV-2 cases at UO using swab-PCR results as a predictor.

<br>


#### Specification

We created a random intercepts logistic regression model with the occurrence of cases (binary) as outcome and swab results for the previous week (the proportion of positive swabs) as predictor. The model has a random intercept for each site.

Our model formula is `cases ~ positives[lag 1 week] + (1|site)`.

The model is fit as follows:

```{r eval=FALSE, echo=TRUE}
swab_model <-
  blme::bglmer(
    cases_binary ~ detection_lag_1week + (1 | site),
    data =  uo_sites,
    family = binomial
  )
```

------------------------------------------------------------------------

```{r model}
swab_model <- read_rds(here('model/uo_swab_model.rds'))
```

<br>

#### Model response time-series

These plots show the swab results, cases, and predictions by the current model.\
<br>

```{r prediction figure live version, fig.height=15, fig.width=8}
# uo cases data - long table
cases_long <-
  read_rds(here('./data/latest_uo_cases.rds')) |>
  filter(!is.na(positive_result),
         positive_result > lubridate::as_date('2021-09-14')) |>
  select(case, role, confirmed_by_test, positive_result, c(`90U`:TBT)) |>
  mutate(
    `Campus Total` = TRUE,
    # 3 days transmissibility
    transmission_start = positive_result - days(3)
    ) |>
  pivot_longer(`90U`:`Campus Total`, names_to = 'site') |>
  filter(value) |>
  select(-value) |>
  mutate(site = fct_relevel(site, "Campus Total", after = 0L))

uo_pred <- read_rds(here('model/uo_pred.rds'))

# create a 3-panel fig for each of 6 buildings.
map(
  # site names for filtering
  .x = swabs$site |> unique() |> str_remove('UO_'),
  .f = ~plot_uo_pred_swab_case(
    swabs = swabs,
    cases = cases_long,
    preds = uo_pred,
    site_select = .x
  )
) |> wrap_plots(guides = 'collect', ncol = 1)

```

<br>

------------------------------------------------------------------------

#### Model summary

These tables show the model coefficients and statistics.

```{r model glance}
broom.mixed::glance(swab_model) |>
  mutate_if(is.numeric, round, 2) |>
  kableExtra::kable(caption = 'Model statistics',
                    format = "html") |>
  kableExtra::kable_styling(bootstrap_options = "striped",
                            full_width = F, position = "left")
```

------------------------------------------------------------------------

```{r model tidy}
broom.mixed::tidy(swab_model) |>
  mutate(p.value = as.character(p.value)) |>
  mutate_if(is.numeric, round, 3) |>
  mutate(p.value = as.numeric(p.value)) |>
  kableExtra::kable(caption = 'Model parameters',
                    format = "html") |>
  kableExtra::kable_styling(bootstrap_options = "striped",
                            full_width = F, position = "left")
# tidy(swab_model, effects = 'ran_vals')
```

<br>

------------------------------------------------------------------------

```{r plot model intercepts}
sjPlot::tab_model(
  swab_model,
  show.re.var = TRUE,
  pred.labels = c("(Intercept)", "Swabs @ t-1week"),
  dv.labels = "Relation between UO cases (y/n) and proportion\n of positive swabs the previous week"
)
```

<br>

------------------------------------------------------------------------

#### Model response curves

This plot shows the modelling data as points (detection lagged 1 week and cases at a given site- y/n), as well as how the probability of future cases varies by the previous weeks detection level and site, according to our model (curves).

```{r response curves, fig.height=4.5}
read_rds(here('fig/plt_model_resp_curves.rds')) +
  labs(subtitle = 'Points show case/swab data and curves show model probability values')
```

<br>

------------------------------------------------------------------------

#### Random Intercepts for Sites

This plot shows the odds ratios for the intercepts of the model (an intercept for each site).

```{r model_sites}
# random effect plot: site intercepts; axes are flipped
sjPlot::plot_model(swab_model, type = 're') +
  theme_light() +
  labs(title = "Random effects", y = 'Odds Ratio')
```

<br>

------------------------------------------------------------------------

## EDA


### Swabs and Cases

This panel shows the cases counts at UO over the course of our sampling period. The case data shown represents the days on which a positive test was reported (black rug lines) and the presumed start of transmissiblity for each case (red lines).

```{r cases and swabs, fig.height=6, fig.width=8, fig.align='center'}
# generated from file R/plot_UO_case_load.R
read_rds(here('fig/plt_cases_swabs.rds'))
```

<br>

------------------------------------------------------------------------

### Swabs, Wifi, and CO~2~

This panel shows linked data from swab results, CO~2~ readings, and wifi traffic (number of users \@ peak, daily) during our study period. Unfortunately, we do not have wifi data for 90U. <br>

```{r uo multivariate panel, fig.width=12.5, fig.asp=0.6}


uo_multivariate <- make_uo_panel(swabs, wifi)
write_rds(uo_multivariate, here('fig/uo_multivariate.rds'))

uo_multivariate & theme(text = element_text(size = 13))
rm(uo_multivariate)
```

<br>

------------------------------------------------------------------------

### Wifi Traffic

```{r wifi_ts, fig.height=6, fig.width=8}
plot_UO_wifi_ts(wifi, swabs)
```

This panel shows a time-series of the daily peak number of wifi users at UO facilities. Sampling days are highlighted in blue.\
<br>

------

<!-- ### Cases: Imputation -->

<!-- I fitted linear models to ascertain the relationship between the positive result date with the symptoms-began date, and/or the isolation-period-end date (3 combinations). I used the three linear models to impute missing result dates by averaging out the predictions of the three models (where data was available). -->
<!-- Data are not missing at random: most of the missing values occur in Jan 2022 as testing became less common and more difficult to access (?). Collection of case records was abandoned by the university altogether in Feb 2022. -->


## Waste Water - Regional

```{r fig.height=3.5, fig.width=5}
regional_ww |>
  group_by(sample_date) |>
  summarise(mean = mean(mean)) |>
  ggplot() +
  geom_rect(data = tibble(xmin = min(swabs$date_swab),
                   xmax = max(swabs$date_swab),
                   ymin = -Inf,
                   ymax = Inf),
            aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
            fill = 'grey90', color = 'gray',
            size = 0.5,
            lty = 2, alpha = 0.5) +
  geom_text(aes(as_date('2022-01-01'), 0.0035, label = 'UOttawa study'),
            color = 'darkgray') +
  geom_point(aes(sample_date, mean),
             alpha = 0.76, size = 0.85, shape = 1) +
  geom_smooth(aes(sample_date, mean),
              method = 'loess', formula = 'y ~ x',
              span = 0.18, size = 0.75, alpha = 0.25,
              fill = 'blue') +
  scale_shape(solid = F) +
  labs(x = 'Date',
       y = 'Mean SARS-CoV-2 / PPMoV ',
       color = 'PCR Target',
       fill = 'PCR Target',
       shape = 'PCR Target',
       subtitle = 'Ottawa Waste-water: Combined N1 and N2 targets')
```



----

<br>

## Time-series comparison

**2022-02-28 Update**: This data uses weekly aggregates, the new results use biweekly aggregates (in case you are wondering why the *r* values aren't the same).


We probably want to summarise our swabs at the weekly (or biweekly?) level for comparison to the waste-water signal. This is what the weekly campus-wide time-series looks like compared with our weekly campus-wide swab positivity:


```{r fig.width = 6, fig.height=6}
ww_weekly_summary <-
  uottawa_ww |>
  mutate(week = get_date_week(sample_date)) |>
  left_join(week_midpoint_date_lookup('2021-01-01', '2022-12-31'),
            by = 'week') |>
  select(date, signal, site) |>
  group_by(date) |>
  summarise(signal_mean = mean(signal),
            signal_sd = sd(signal))

plt_ww_weekly <- ww_weekly_summary |>
  ggplot(aes(date,
             signal_mean,
             # ymin = signal_mean - signal_sd,
             # ymax = signal_sd + signal_sd
             )) +
  geom_line() +
  geom_point() +
  geom_point(data = uottawa_ww,
             aes(as_date(sample_date), signal),
             size = 1, shape = 1, alpha = 0.5) +
  labs(x = 'Date', y = 'SARS-CoV-2 / PMMoV',
       title = 'Weekly campus-wide average SARS-CoV-2 waste-water detection')

swabs_weekly_summary <- swabs |>
  mutate(week = get_date_week(date_swab)) |>
  left_join(week_midpoint_date_lookup('2021-01-01', '2022-12-31'),
            by = 'week') |>
  group_by(date) |>
  summarise(
    x = sum(pcr_positive == 'Positive'),
    n = n(),
    positivity = list(binom_ci(x, n))
    ) |>
  unnest_wider(positivity)

plt_swabs_weekly <- swabs_weekly_summary |>
  ggplot(aes(x = date, y = point_est*100)) +
  geom_line() +
  geom_point() +
  scale_x_date() +
  scale_y_continuous(limits = c(0,100)) +
  labs(x = 'Date', y = 'Swab positivity (%)',
       title = 'Weekly campus-wide surface swab detection rate')

wrap_plots(plt_swabs_weekly, plt_ww_weekly, ncol = 1)
```

If we look at the relationship between swabs and waste-water on weeks where both results are available, we see that they have a positive correlation.

```{r fig.width = 5, fig.height=5}
campus_ww_swab_summary <-
  swabs_weekly_summary |>
  select(date, point_est) |>
  left_join(ww_weekly_summary, by = 'date') |>
  transmute(date, ww = signal_mean, swabs = point_est*100)

campus_ww_swab_summary |>
  ggplot(aes(swabs, ww, color = date)) +
  geom_smooth(method = 'lm', color = 'gray', fill = 'gray') +
  geom_point() +
  scale_color_date() +
  coord_fixed(ratio = 18000) +
  labs(x = 'Swab positivity (%)',
       y = 'Waste-water signal',
       color = "Sample\nWeek")
```


```{r}
cortest_swab_ww <-
  lst(
    pearson = cor.test(
      campus_ww_swab_summary$ww,
      campus_ww_swab_summary$swabs,
      method = 'pearson'
    ),
    pearson_est = pearson$estimate |> round(2),
    pearson_lower = pearson$conf.int[1] |> round(2),
    pearson_upper = pearson$conf.int[2] |> round(2),

    spearman = cor.test(
      campus_ww_swab_summary$ww,
      campus_ww_swab_summary$swabs,
      method = 'spearman'
    ),
    spearman_est = spearman$estimate |> round(3),
    # spearman_lower = spearman$conf.int[1] |> round(2),
    # spearman_upper = spearman$conf.int[2] |> round(2),

    kendall = cor.test(
      campus_ww_swab_summary$ww,
      campus_ww_swab_summary$swabs,
      method = 'kendall'
  ),
    kendall_est = kendall$estimate |> round(2),
    # kendall_lower = kendall$conf.int[1] |> round(2),
    # kendall_upper = kendall$conf.int[2] |> round(2),

) |>
  suppressWarnings() |>
  glue_data(
    "The Pearson's *r* (linear correlation) is {pearson_est} ({pearson_lower}-{pearson_upper}).
    The Spearman's *r* (rank correlation) is {spearman_est}..
    The Kendall's *Tau*  is {kendall_est}."
  )

```


Pearson's *r* is more affected by outliers. Spearman's *r* or Kendall's *Tau* are probably the more reliable measures of correlation between these variables.

<br>
<br>

(Data doesn't appear normally distributed, btw)

```{r fig.height=3, fig.width=7.5}
campus_ww_swab_summary |>
  pivot_longer(cols = c(ww, swabs)) |>
  ggplot(aes(value)) +
  geom_histogram(bins = 15, color = 'lightgray') +
  facet_wrap(~name, scales = 'free')
```

<br>
<br>

-----

## Waste Water - UO

I've removed the off-campus site (Robert Guidon Hall) and retained the other 6 near-campus sites (raw data below).

```{r fig.width=6, fig.height=5}
uottawa_ww |>
  ggplot(aes(sample_date, signal)) +
  geom_path(alpha = 0.6) +
  geom_point(alpha = 0.5, size = 0.5) +
  labs(x = 'Date', y = 'SARS-CoV-2 / PPMoV') +
  facet_wrap(~site, ncol = 2)

```


