---
title: "CUBE - UOttawa"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_chunks: hide
    toc: true
    toc_float: true
    theme: spacelab
---

```{r setup, echo=F}
knitr::opts_chunk$set(echo = F, message = F, warning = F)

```


```{r}
library(tidyverse)
library(here)
library(ggiraph)
library(patchwork)
ggplot2::theme_set(theme_bw())
source(here('R/report_plots.R'))
```


```{r}
# swab data
swabs <- read_rds(here('data/cube.rds')) |> 
  filter(str_detect(site, '^UO_')) 

# waste water data 
raw_ww <-
  readxl::read_excel(here::here('data/wastewater.xlsx')) |>
  janitor::clean_names()

uottawa_ww <-
  raw_ww |>
  filter(!is.na(sample_date)) |>
  mutate(
    source =
      source_name |>
      str_remove('^Data - uOttawa - ') |>
      str_remove('.xlsx$'),
    site = case_when(
      source == 'uo_aa' ~ 'AA',
      source == 'uo_fa' ~ 'FA',
      source == 'uo_ft' ~ 'FT',
      source == 'uo_na' ~ 'NA',
      source == 'uo_rgn' ~ 'RGN',
      source == 'uo_st' ~ 'ST',
      source == 'uo_tbt' ~ 'TBT',
      TRUE ~ 'not matched'
    )
  )

# ottawa wastewater data: summary
summarise_wastewater <- function(){
  read_rds(here('data/ww_ottawa.rds')) |> 
  select(sample_date, starts_with('cov_')) |> 
  pivot_longer(contains('cov_')) |> 
  mutate(target = str_extract(name, 'cov_n.'),
         stat = str_extract(name, 'mean|sd'),
         ) |> 
  select(-name) |> 
  pivot_wider(names_from = stat, values_from = value) |> 
  mutate(.lower = mean - sd, .upper = mean + sd)   
}

wifi <- 
  read_rds(here('data/uo_wifi.rds')) |> 
  filter(date >= min(swabs$date_swab), date <= max(swabs$date_swab))

```



## 📊 Summary

This plot shows the counts of positive (red) and negative (yellow) samples collected at each facility over time (x-axis). Samples that could not be tested are shown in navy. Only flocked swabs are shown. (Other sponge swabs were collected on 2022-04-28 were for comparing flocked and sponge swabs.)

```{r university bargraph}
plot_timeseries_summary(
  swabs |> 
    filter(
      swab_type %in% 'flock',
      !negative_control,
      !is.na(negative_control)
    ),
  height_svg = 3.3,
  width_svg = 8
  )
```


------------------------------------------------------------------------

## ⡯⡷ Dotplot

This plot shows the counts of positive and negative samples collected at each facility over time. 

```{r univeristy dotplot}
plot_timeseries_locations(
  swabs |> 
    filter(swab_type == 'flock'),
  height_svg = 6,
  width_svg = 8
  )
```

------------------------------------------------------------------------

## Waste Water - UO


```{r}
uottawa_ww |>
  ggplot(aes(sample_date, viral_copies_per_copies_pep_avg, color = site)) +
  geom_line(alpha = 0.5) +
  geom_point(alpha = 0.5) +
  labs(x = 'Date', y = 'SARS-CoV-2 / PPMoV')

rm(raw_ww, uottawa_ww)
```



-----



## Model {.tabset}

This section contains results from modeling SARS-CoV-2 cases at UO using swab-PCR results as a predictor.

<br>


#### Specification

We created a random intercepts logistic regression model with the occurrence of cases (binary) as outcome and swab results for the previous week (the proportion of positive swabs) as predictor. The model has a random intercept for each site.

Our model formula is `cases ~ positives[lag 1 week] + (1|site)`.

The model is fit as follows:

```{r eval=FALSE, echo=TRUE}
swab_model <-
  blme::bglmer(
    cases_binary ~ detection_lag_1week + (1 | site),
    data =  uo_sites,
    family = binomial
  )
```

------------------------------------------------------------------------

```{r model}
swab_model <- read_rds(here('model/uo_swab_model.rds'))
```

<br>

#### Model response time-series

These plots show the swab results, cases, and predictions by the current model.\
<br>

```{r prediction figure live version, fig.height=15, fig.width=8}
# uo cases data - long table
cases_long <-
  read_rds(here('./data/latest_uo_cases.rds')) |>
  filter(!is.na(positive_result),
         positive_result > lubridate::as_date('2021-09-14')) |>
  select(case, role, confirmed_by_test, positive_result, c(`90U`:TBT)) |>
  mutate(
    `Campus Total` = TRUE,
    # 3 days transmissibility
    transmission_start = positive_result - days(3)
    ) |>
  pivot_longer(`90U`:`Campus Total`, names_to = 'site') |>
  filter(value) |>
  select(-value) |> 
  mutate(site = fct_relevel(site, "Campus Total", after = 0L))

uo_pred <- read_rds(here('model/uo_pred.rds'))

# create a 3-panel fig for each of 6 buildings.
map(
  # site names for filtering 
  .x = swabs$site |> unique() |> str_remove('UO_'),
  .f = ~plot_uo_pred_swab_case(
    swabs = swabs,
    cases = cases_long,
    preds = uo_pred,
    site_select = .x
  )
) |> wrap_plots(guides = 'collect', ncol = 1)

```

<br>

------------------------------------------------------------------------

#### Model summary

These tables show the model coefficients and statistics.

```{r model glance}
broom.mixed::glance(swab_model) |> 
  mutate_if(is.numeric, round, 2) |> 
  kableExtra::kable(caption = 'Model statistics',
                    format = "html") |> 
  kableExtra::kable_styling(bootstrap_options = "striped", 
                            full_width = F, position = "left")
```

------------------------------------------------------------------------

```{r model tidy}
broom.mixed::tidy(swab_model) |>
  mutate(p.value = as.character(p.value)) |> 
  mutate_if(is.numeric, round, 3) |> 
  mutate(p.value = as.numeric(p.value)) |> 
  kableExtra::kable(caption = 'Model parameters',     
                    format = "html") |> 
  kableExtra::kable_styling(bootstrap_options = "striped",
                            full_width = F, position = "left")
# tidy(swab_model, effects = 'ran_vals')
```

<br>

------------------------------------------------------------------------

```{r plot model intercepts}
sjPlot::tab_model(
  swab_model,
  show.re.var = TRUE,
  pred.labels = c("(Intercept)", "Swabs @ t-1week"),
  dv.labels = "Relation between UO cases (y/n) and proportion\n of positive swabs the previous week"
)
```

<br>

------------------------------------------------------------------------

#### Model response curves

This plot shows the modelling data as points (detection lagged 1 week and cases at a given site- y/n), as well as how the probability of future cases varies by the previous weeks detection level and site, according to our model (curves).

```{r response curves, fig.height=4.5}
read_rds(here('fig/plt_model_resp_curves.rds')) +
  labs(subtitle = 'Points show case/swab data and curves show model probability values')
```

<br>

------------------------------------------------------------------------

#### Random Intercepts for Sites

This plot shows the odds ratios for the intercepts of the model (an intercept for each site).

```{r model_sites}
# random effect plot: site intercepts; axes are flipped
sjPlot::plot_model(swab_model, type = 're') +
  theme_light() +
  labs(title = "Random effects", y = 'Odds Ratio')
```

<br>

------------------------------------------------------------------------

### Swabs and Cases

This panel shows the cases counts at UO over the course of our sampling period. The case data shown represents the days on which a positive test was reported (black rug lines) and the presumed start of transmissiblity for each case (red lines).

```{r cases and swabs, fig.height=6, fig.width=8, fig.align='center'}
# generated from file R/plot_UO_case_load.R
read_rds(here('fig/plt_cases_swabs.rds'))
```

<br>

------------------------------------------------------------------------

### Swabs, Wifi, and CO~2~

This panel shows linked data from swab results, CO~2~ readings, and wifi traffic (number of users \@ peak, daily) during our study period. Unfortunately, we do not have wifi data for 90U. <br>

```{r uo multivariate panel, fig.width=12.5, fig.asp=0.6}
uo_multivariate <- make_uo_panel(swabs, wifi)
write_rds(uo_multivariate, here('fig/uo_multivariate.rds'))

uo_multivariate & theme(text = element_text(size = 13))
rm(uo_multivariate)
```

<br>

------------------------------------------------------------------------

### Wifi Traffic

```{r wifi_ts, fig.height=6, fig.width=8}

plot_UO_wifi_ts(wifi, swabs)

```

This panel shows a time-series of the daily peak number of wifi users at UO facilities. Sampling days are highlighted in blue.\
<br>

### 